Supplementary Information for RVD2: An ultra-sensitive variant detection model for low-depth targeted next-generation sequencing data

### 2014-03-14 doc/bioinformatics2014/document.tex
Update the manuscript and supplementary information

\documentclass{bioinfo}
\copyrightyear{2014}
\pubyear{2014}
\usepackage{amsmath}
\usepackage{afterpage}
\usepackage{algorithm}
\usepackage{bbm}

@@

\section{Results:}
We present a novel variant calling algorithm that uses a hierarchical Bayesian model to estimate allele frequency and call variants in heterogeneous samples. We show that our algorithm improves upon current classifiers and has higher sensitivity and specificity over a wide range of median read depth and minor allele frequency. We identify six mutations in the PAXP1 gene in a matched clinical breast ductal carcinoma tumor sample; two of which are loss-of-heterozygosity events.

@@

We used RVD2 to identify germline and somatic mutations in the diploid HCC1187 sample. To identify germline mutations, we used the control data in the following hypothesis test $\Pr( \tilde{\mu}_j^{\text{control}} \in [\tau_\text{l},\tau_\text{u}] ) > 1-\alpha$, where the intervals are: homozygous reference $[0, 0.2]$, heterozygous mutant $[0.1, 0.9]$, and homozygous mutant $[0.8, 1.0]$ and the size of the test is $\alpha=0.05$.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{pdf_figs/ROC_without_chi2.pdf}
\caption{ROC curve varying read depth showing detection performance.}
\label{fig:ROC}
\end{center}
\end{figure}

@@

\begin{figure*}[!tpb]
\begin{center}
\includegraphics[width=0.9\textwidth]{pdf_figs/HCC1187_MuBarPlot.pdf}
\caption{Estimated minor allele fraction for called variants in PAXIP1 gene. Star (*) indicates $\mu^{\text{case}}$  is significantly different from $ \mu^{\text{control}}$.}
\label{fig:brca_MAF}
\end{center}
\end{figure*}
%('rvd2\results\2013-11-21_HCC1187_PAXIP1_mu_barplot_T0')
%TODO: add star to test that have mudelta > .16 Done.
%TODO: set tau=0.16 for mudelta test.  Done

>>>>

\documentclass{bioinfo}
\copyrightyear{2014}
\pubyear{2014}
\usepackage{amsmath}
\usepackage{afterpage}
\usepackage{algorithm}
\usepackage{bbm}
\usepackage{subfig}

@@

\section{Results:}

%TODO Update the chinical dataset result
We present a novel variant calling algorithm that uses a hierarchical Bayesian model to estimate allele frequency and call variants in heterogeneous samples. We show that our algorithm improves upon current classifiers and has higher sensitivity and specificity over a wide range of median read depth and minor allele frequency. We identify six mutations in the PAXP1 gene in a matched clinical breast ductal carcinoma tumor sample; two of which are loss-of-heterozygosity events.

@@

%TODO: change the result
We used RVD2 to identify germline and somatic mutations in the diploid HCC1187 sample. To identify germline mutations, we used the control data in the following hypothesis test $\Pr( \tilde{\mu}_j^{\text{control}} \in [\tau_\text{l},\tau_\text{u}] ) > 1-\alpha$, where the intervals are: homozygous reference $[0, 0.05)$, heterozygous mutant $[0.5, 0.0.75)$, and homozygous mutant $[0.75, 1.0]$.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.45\textwidth]{pdf_figs/ROC_without_chi2.pdf}
\caption{ROC curve varying read depth showing detection performance.}
\label{fig:ROC}
\end{center}
\end{figure}

@@

\begin{figure*}[!tpb]
\begin{center}
\includegraphics[width=0.7\textwidth]{pdf_figs/HCC1187_MuBarPlot.pdf}
\caption{Estimated minor allele fraction for called variants in PAXIP1 gene. Star (*) indicates $\mu^{\text{case}}$  is significantly different from $ \mu^{\text{control}}$.}
\label{fig:brca_MAF}
\end{center}
\end{figure*}
%('rvd2\results\2013-11-21_HCC1187_PAXIP1_mu_barplot_T0')
%TODO: add star to test that have mudelta > .16 Done.
%TODO: set tau=0.16 for mudelta test.  Done

\begin{figure*}[!tpb]
    \centering
    \subfloat[Estimated minor allele fraction for called variants in PAXIP1 gene.  $ \mu^{\text{control}}$.]{{\includegraphics[width=0.8\textwidth]{pdf_figs/HCC1187_dotplot.pdf} }}%
    %
    \subfloat[Depth Chart for Positions called by RVD2 and muTect]{{\includegraphics[width=0.2\textwidth]{pdf_figs/position_lookup_chart.pdf} }}%
    \caption{2 Figures side by side}%
    \label{fig:example}%
\end{figure*}

### 2014-03-14 doc/bioinformatics2014/supplementary_info.tex
Update the manuscript and supplementary information

\documentclass[11pt,reqno]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{algorithm, algorithmic}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{epstopdf}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

@@

\section{Histogram of $ \mu_j $ for mutations detected by RVD2 in HCC1187 dataset}
\begin{figure}[H]
\begin{center}
\includegraphics[width=120mm]{pdf_figs/HCC1187_histogram_neg.pdf}
\caption{Estimated minor allele fraction for called variants in 1.0\% dilution.}
\label{fig:hist_neg}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=120mm]{pdf_figs/HCC1187_histogram_pos.pdf}
\caption{Estimated minor allele fraction for called variants in 1.0\% dilution.}
\label{fig:hist_pos}
\end{center}
\end{figure}

\section{Comparison between RVD2, VarScan2 somatic and muTect in Clincal HCC1187 dataset}
\begin{figure}[H]
\begin{center}
\includegraphics[width=1\textwidth]{pdf_figs/HCC1187_RVD2.pdf}
\caption{RVD2.}
\label{fig:rd_RVD2}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.7\textwidth]{pdf_figs/HCC1187_mutect.pdf}
\caption{muTect.}
\label{fig:rd_mutect}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.8\textwidth]{pdf_figs/HCC1187_varscan2_true_positives.pdf}
\caption{HCC1187 varscan2 true positives.}
\label{fig:rd_varscan2_TP}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.8\textwidth]{pdf_figs/HCC1187_varscan2_false_negatives.pdf}
\caption{ HCC1187 varscan2 false negatives.}
\label{fig:rd_varscan2_FN}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.8\textwidth]{pdf_figs/HCC1187_varscan2_false_positives.pdf}
\caption{HCC1187 varscan2 false positives}
\label{fig:rd_varscan2_FP}
\end{center}
\end{figure}

>>>>

\documentclass[11pt,reqno]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{algorithm, algorithmic}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pdfpages}
%\usepackage{epstopdf}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

@@

\section{Histogram of $ \mu_j $ for mutations detected by RVD2 in HCC1187 dataset}
\begin{figure}[H]
\begin{center}
\includegraphics[width=120mm]{pdf_figs/HCC1187_histogram_pos.pdf}
\caption{Estimated minor allele fraction for called variants in 1.0\% dilution.}
\label{fig:hist_pos}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=120mm]{pdf_figs/HCC1187_histogram_neg.pdf}
\caption{Estimated minor allele fraction for called variants in 1.0\% dilution.}
\label{fig:hist_neg}
\end{center}
\end{figure}

\section{Comparison between RVD2, VarScan2 somatic and muTect in Clincal HCC1187 dataset}
\includepdf[pages={1-},scale=0.9]{pdf_figs/position_lookup_chart_all.pdf}
